### 2019-12-11 ic-journal/expansion.tex
Up to Section 4.0
Up to Section 4.2

\emph{Strong} bisimulation and strong bisimilarity ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in 
clause (1) 
 with the transition $Q \arr \mu Q'$ (and similarly for the other clause).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), i.e.~$P_1 \wb Q_1$ and $P_2 \wb Q_2$  \hl{do} not imply
 $P_1 + P_2 \wb Q_1 + Q_2$.
For this, Milner introduced the concept of \emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, if % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}
% This new sentence maintains a link with Definition 2.1
Relation $\rapprox$ is indeed preserved by the sum operator.
% \hl{It can be proved that $P_1 \rapprox Q_1 \wedge P_2 \rapprox Q_2 \Longrightarrow
% P_1 + P_2 \rapprox Q_1 + Q_2$.}
The above definition also brings up a proof technique for proving rooted
bisimilarity from a given bisimulation. The next
lemma plays an
important role in proving some key results in this paper:
\begin{lemma}[Rooted bisimilarity by bisimulation]
\label{l:obsCongrByWeakBisim}
Given a (weak) bisimulation $\RR$, suppose two processes $P$ and $Q$
satisfy the following properties:
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \RR Q'$.
\end{enumerate}
Then $P$ and $Q$ are rooted bisimilar, i.e.~$P \approx^c Q$.
\end{lemma}

>>>>

\emph{Strong} bisimulation and strong bisimilarity ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in 
clause (1) 
 with the transition $Q \arr \mu Q'$ (and similarly for the other clause).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), i.e.~$P_1 \wb Q_1$ and $P_2 \wb Q_2$  \hl{do} not imply
 $P_1 + P_2 \wb Q_1 + Q_2$.
For this \hl{reason}, Milner introduced the concept of
\emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, if % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}
% This new sentence maintains a link with Definition 2.1
Relation $\rapprox$ is indeed preserved by the sum operator.
% \hl{It can be proved that $P_1 \rapprox Q_1 \wedge P_2 \rapprox Q_2 \Longrightarrow
% P_1 + P_2 \rapprox Q_1 + Q_2$.}
The above definition also brings up a proof technique for proving rooted
bisimilarity from a given bisimulation. The next
lemma plays an
important role in proving some key results in this paper:
\begin{lemma}[rooted bisimilarity by bisimulation]
\label{l:obsCongrByWeakBisim}
Given a (weak) bisimulation $\RR$, suppose two processes $P$ and $Q$
satisfy the following properties:
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \RR Q'$.
\end{enumerate}
Then $P$ and $Q$ are rooted bisimilar, i.e.~$P \approx^c Q$.
\end{lemma}

### 2019-12-11 ic-journal/formal.tex
Up to Section 4.2

Higher Order Logic (HOL)~\cite{hollogic} traces
its roots back to the
\emph{Logic of Computable Functions
  (LCF)}~\cite{gordon1979edinburgh,milner1972logic}
by Robin Milner and others
since 1972. It is a variant of
Church’s Simple Theory of Types (STT)~\cite{church1940formulation},
plus a higher order version of Hilbert's choice operator $\varepsilon$,
Axiom of Infinity, and Rank-1 (prenex) polymorphism.
HOL4 has implemented the original HOL, 
while some other theorem provers in the HOL family (e.g.~Isabelle/HOL) have
certain extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed, HOL has considerably simpler logical
foundations than most other theorem provers.
As a consequence, theories and proofs verified in HOL are easier to understand
 \hl{for} people who are not familar with more advanced
dependent type theories, \hl{e.g.~those implemented by Coq.}

HOL4 is implemented in Standard ML, and
 the same programming language
plays three different roles:
\begin{enumerate}
\item The underlying implementation language for the core HOL engine;
\item The language in which proof tactics are implemented;
\item The interface language of the HOL proof scripts and interactive shell.
\end{enumerate}
Moreover, using the same language HOL4 users can write complex automatic
verification tools by calling HOL's theorem proving
facilities. The formal proofs of the CCS theorems that we have carried
out 
are mostly done in a manner that closely follows
their paper proofs, with minimal automatic proof searching.

@@

HOL is a formal system of typed logical terms. The types are expressions that denote sets (in the
universe $\mathcal{U}$). HOL type system is much simpler than those
based on dependent types and other type theories. There are four kinds of types in the HOL
logic, as illustrated in Fig. \ref{fig:hol-types} for its BNF
grammar. In HOL, the standard atomic types \emph{bool} and \emph{ind}
 denote, respectively, the distinguished two-element set 2 and the
distinguished infinite set $I$.

@@

The terms of the HOL logic represent elements of the
sets denoted by the types. There are four kinds of terms in the HOL
logic. They can be described (in simplified forms) by the BNF grammar in
Fig.~\ref{fig:hol-terms}; see~\cite{hollogic} for a complete description.

@@

The deductive system of HOL is specified by eight primitive
derivative rules~\cite{hollogic}:
\begin{enumerate}
\item Assumption introduction (\texttt{ASSUME});
\item Reflexivity (\texttt{REFL});
\item $\beta$-conversion (\texttt{BETA\_CONV});
\item Substitution (\texttt{SUBST});
\item Abstraction (\texttt{ABS});
\item Type instantiation (\texttt{INST\_TYPE});
\item Discharging an assumption (\texttt{DISCH});
\item Modus Ponens (\texttt{MP}).
\end{enumerate}
All proofs are eventually reduced to applications of the above rules,
which also give the semantics of two foundamental
logical connectives,  equality ($=$) and implication
($\Rightarrow$). The remaining logical connectives and first-order
quantifiers, including the logical true (\HOLinline{\HOLConst{T}}) and false (\HOLinline{\HOLConst{F}}), are
further defined as $\lambda$-functions: 
\begin{equation*}
\begin{array}{l}
\turn \T \HOLTokenDefEquality ((\lquant{x_{\ty{bool}}}x) =
               (\lquant{x_{\ty{bool}}}x))    \\
\turn \forall \HOLTokenDefEquality \lquant{P_{\alpha\fun\ty{bool}}} P =
                    (\lquant{x}\T ) \\
\turn \exists \HOLTokenDefEquality \lquant{P_{\alpha\fun\ty{bool}}} P({\hilbert}\, P) \\
\turn \F \HOLTokenDefEquality \uquant{b_{\ty{bool}}} b  \\
\turn \neg \HOLTokenDefEquality \lquant{b} b \imp \F \\
\turn {\wedge}  \HOLTokenDefEquality  \lquant{b_1\ b_2}\uquant{b} (b_1\imp (b_2 \imp b)) \imp b \\
\turn {\vee} \HOLTokenDefEquality \lquant{b_1\ b_2}\uquant{b} (b_1 \imp b)\imp ((b_2 \imp b) \imp b) \\
\turn \OneOne \HOLTokenDefEquality \lquant{f_{\alpha \fun\beta}}\uquant{x_1\ x_2}
                    (f\ x_1 = f\ x_2)  \imp (x_1 = x_2) \\
\turn \Onto \HOLTokenDefEquality \lquant{f_{\alpha\fun\beta}}
                  \uquant{y}\equant{x} y = f\ x \\
\turn \TyDef \HOLTokenDefEquality \lambda P_{\alpha\fun\ty{bool}}\
                  rep_{\beta\fun\alpha}.\;
                  \OneOne\ rep \ \wedge{}\  (\uquant{x}P x = (\equant{y} x = rep\ y))
\end{array}
\end{equation*}
The last logical constant,
$\TyDef$, can be used to define new HOL types as bijections of
subsets of existing types~\cite{Melham:1989dk}.
HOL \texttt{Datatype} package~\cite{Melham:1991, holdesc} automates
this tedious process, and can be used for defining the 
types needed for CCS.
%
Finally, the whole HOL \emph{standard} theory is based on the
following four axioms:\footnote{HOL is strictly
  weaker than ZFC (the Zermelo-Frankel set theory with the
Axiom of Choice), thus not all theorems valid in ZFC can be formalised
in HOL. (See~\cite{hollogic} for more details.)}
\begin{equation*}
\begin{array}{@{}l@{\qquad}l}
\mbox{\texttt{BOOL\_CASES\_AX}} &\vdash \uquant{b} (b = \T )\ \vee \ (b = \F )\\
\mbox{\texttt{ETA\_AX}} &
\vdash \uquant{f_{\alpha\fun\beta}}(\lquant{x}f\ x) = f\\
\mbox{\texttt{SELECT\_AX}} &
\vdash \uquant{P_{\alpha\fun\ty{bool}}\ x} P\ x \imp P({\hilbert}\ P)\\
\mbox{\texttt{INFINITY\_AX}}&
\vdash \equant{f_{\ind\fun \ind}} \OneOne \ f \conj \neg(\Onto \ f)\\
\end{array}
\end{equation*}

>>>>

Higher Order Logic (HOL)~\cite{hollogic} traces
its roots back to the
\emph{Logic of Computable Functions
  (LCF)}~\cite{gordon1979edinburgh,milner1972logic}
by Robin Milner and others
since 1972. It is a variant of
Church’s Simple Theory of Types (STT)~\cite{church1940formulation},
plus a higher order version of Hilbert's choice operator $\varepsilon$,
Axiom of Infinity, and Rank-1 (prenex) polymorphism.
HOL4 has implemented the original HOL, 
while some other theorem provers in the HOL family (e.g.~Isabelle/HOL) have
certain extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed, HOL has considerably simpler logical
foundations than most other theorem provers.
As a consequence, theories and proofs verified in HOL are easier to understand
 \hl{for} people who are not familar with more advanced dependent type
 theories, \hl{e.g. the Calculus of Constructions implemented in Coq.}

HOL4 is implemented in Standard ML, and
 the same programming language
plays three different roles:
\begin{itemize}
\item The underlying implementation language for the core HOL engine;
\item The language in which proof tactics are implemented;
\item The interface language of the HOL proof scripts and interactive shell.
\end{itemize}
Moreover, using the same language HOL4 users can write complex automatic
verification tools by calling HOL's theorem proving
facilities. The formal proofs of the CCS theorems that we have carried
out 
are mostly done in a manner that closely follows
their paper proofs, with minimal automatic proof searching.

@@

HOL is a formal system of typed logical terms. The types are expressions that denote sets (in the
universe $\mathcal{U}$). HOL type system is much simpler than those
based on dependent types and other type theories. There are four kinds of types in the HOL
logic, as illustrated in Fig. \ref{fig:hol-types} for its BNF
grammar. In HOL, the standard atomic types \emph{bool} and \emph{ind}
 denote, respectively, the distinguished two-element set $\mathbf{2}$ and the
distinguished infinite set $\mathbf{I}$.

@@

\hl{HOL terms} represent elements of the
sets denoted by \hl{their} types. There are four kinds of \hl{HOL terms,
which} can be described (in simplified forms) by the BNF grammar in
Fig.~\ref{fig:hol-terms}. (See~\cite{hollogic} for a complete
description of HOL, \hl{including the primitive derivative rules to be
mentioned below.})

@@

The deductive system of HOL is specified by eight primitive
derivative rules:
\begin{enumerate}
\item Assumption introduction (\texttt{ASSUME});
\item Reflexivity (\texttt{REFL});
\item $\beta$-conversion (\texttt{BETA\_CONV});
\item Substitution (\texttt{SUBST});
\item Abstraction (\texttt{ABS});
\item Type instantiation (\texttt{INST\_TYPE});
\item Discharging an assumption (\texttt{DISCH});
\item Modus Ponens (\texttt{MP}).
\end{enumerate}
All proofs are eventually reduced to applications of the above rules,
which also give the semantics of two foundamental
logical connectives,  equality ($=$) and implication
($\Rightarrow$). The remaining logical connectives and first-order
quantifiers, including the logical true (\HOLinline{\HOLConst{T}}) and false (\HOLinline{\HOLConst{F}}), are
further defined as $\lambda$-functions: 
\begin{equation*}
\begin{array}{l}
\turn \T \HOLTokenDefEquality ((\lquant{x_{\ty{bool}}}x) =
               (\lquant{x_{\ty{bool}}}x))    \\
\turn \forall \HOLTokenDefEquality \lquant{P_{\alpha\fun\ty{bool}}} P =
                    (\lquant{x}\T ) \\
\turn \exists \HOLTokenDefEquality \lquant{P_{\alpha\fun\ty{bool}}} P({\hilbert}\, P) \\
\turn \F \HOLTokenDefEquality \uquant{b_{\ty{bool}}} b  \\
\turn \neg \HOLTokenDefEquality \lquant{b} b \imp \F \\
\turn {\wedge}  \HOLTokenDefEquality  \lquant{b_1\ b_2}\uquant{b} (b_1\imp (b_2 \imp b)) \imp b \\
\turn {\vee} \HOLTokenDefEquality \lquant{b_1\ b_2}\uquant{b} (b_1 \imp b)\imp ((b_2 \imp b) \imp b) \\
\turn \OneOne \HOLTokenDefEquality \lquant{f_{\alpha \fun\beta}}\uquant{x_1\ x_2}
                    (f\ x_1 = f\ x_2)  \imp (x_1 = x_2) \\
\turn \Onto \HOLTokenDefEquality \lquant{f_{\alpha\fun\beta}}
                  \uquant{y}\equant{x} y = f\ x \\
\turn \TyDef \HOLTokenDefEquality \lambda P_{\alpha\fun\ty{bool}}\
                  rep_{\beta\fun\alpha}.\;
                  \OneOne\ rep \ \wedge{}\  (\uquant{x}P x = (\equant{y} x = rep\ y))
\end{array}
\end{equation*}
The last logical constant,
$\TyDef$, can be used to define new HOL types as bijections of
subsets of existing types~\cite{Melham:1989dk}.
HOL \texttt{Datatype} package~\cite{Melham:1991, holdesc} automates
this tedious process, and can be used for defining the 
types needed for CCS.
%
Finally, the whole HOL \emph{standard} theory is based on the
following four axioms:\footnote{HOL is strictly
  weaker than ZFC (the Zermelo-Frankel set theory with the
Axiom of Choice), thus not all theorems valid in ZFC can be formalised
in HOL. (See~\cite{hollogic} for more details.)}
\begin{equation*}
\begin{array}{@{}l@{\qquad}l}
\mbox{\texttt{BOOL\_CASES\_AX}} &\vdash \uquant{b} (b = \T )\ \vee \ (b = \F )\\
\mbox{\texttt{ETA\_AX}} &
\vdash \uquant{f_{\alpha\fun\beta}}(\lquant{x}f\ x) = f\\
\mbox{\texttt{SELECT\_AX}} &
\vdash \uquant{P_{\alpha\fun\ty{bool}}\ x} P\ x \imp P({\hilbert}\ P)\\
\mbox{\texttt{INFINITY\_AX}}&
\vdash \equant{f_{\ind\fun \ind}} \OneOne \ f \conj \neg(\Onto \ f)\\
\end{array}
\end{equation*}
